---
title: "Group Project"
author: "Evan McKee, Roman, Jordan"
date: "12/06/2022"
output: pdf_document
---


## Input data
```{r, message=FALSE}
## setting working directory
setwd("C:/git/statsGroupProject/R")
## input data
load("ces.RData")

library(ggplot2)
```


## Variables to factor
```{r, message=FALSE}

data <- load("ces.RData")


ces$polengage <- ces$voted + ces$meeting + ces$sign +
    ces$campaign + ces$protest + ces$contact + ces$donate


## Region
ces$region <- factor(ces$region,
    levels = 1:4,
    labels = c("Northeast", "Midwest", "South", "West")
)

## Gender
ces$gender <- factor(ces$gender,
    levels = 1:2,
    labels = c("Male", "Female")
)

## Education
ces$educ <- factor(ces$educ,
    levels = 1:6,
    labels = c(
        "No HS", "High School Graduate", "Some college",
        "2-year degree", "4-year degree", "Post-grad"
    )
)

## Race
ces$race <- factor(ces$race,
    levels = 1:8,
    labels = c(
        "White", "Black", "Hispanic", "Asian",
        "Native American", "Middle Eastern", "Two or more races", "Other"
    )
)

## Hispanic
ces$hispanic <- factor(ces$hispanic,
    levels = 1:2,
    labels = c("Yes", "No")
)

## Homeowner
ces$ownhome <- factor(ces$ownhome,
    levels = 1:3,
    labels = c("Own", "Rent", "Other")
)

## Neighborhood Type
ces$urbancity <- factor(ces$urbancity,
    levels = 1:5,
    labels = c("City", "Suburb", "Town", "Rural Area", "Other")
)

## Union
ces$unionhh <- factor(ces$unionhh,
    levels = 1:2,
    labels = c("Yes", "No")
)

## Religious
ces$religious <- factor(ces$religious,
    levels = 1:4,
    labels = c(
        "Very important", "Not too important",
        "Not too important", "Not at all important"
    )
)

## Family income
ces$faminc2 <- factor(ces$faminc,
    levels = 1:16,
    labels = c(
        "< 10,000", "10,000 - 19,999", "20,000 - 29,999",
        "30,000 - 39,999", "40,000 - 49,999",
        "50,000 - 59,999", "60,000 - 69,999", "70,000 - 79,999",
        "80,000 - 99,999", "100,000-119,000", "120,000 - 149,999",
        "150,000 - 199,999", "200,000 - 249,999", "250,000 - 349,999",
        "350,000 - 499,999", "500,000 or more"
    )
)

## Voted
ces$voted <- factor(ces$voted,
    levels = 1:2,
    labels = c("Yes", "No")
)

## Meeting
ces$meeting <- factor(ces$meeting,
    levels = 1:2,
    labels = c("Yes", "No")
)

## Signed
ces$sign <- factor(ces$sign,
    levels = 1:2,
    labels = c("Yes", "No")
)

## Campaign
ces$campaign <- factor(ces$campaign,
    levels = 1:2,
    labels = c("Yes", "No")
)

## Protest
ces$protest <- factor(ces$protest,
    levels = 1:2,
    labels = c("Yes", "No")
)

## Contact
ces$contact <- factor(ces$contact,
    levels = 1:2,
    labels = c("Yes", "No")
)

## Donate
ces$donate <- factor(ces$donate,
    levels = 1:2,
    labels = c("Yes", "No")
)

## ideo5
ces$ideo5 <- factor(ces$ideo5,
    levels = 1:5,
    labels = c(
        "Very liberal", "Liberal", "Moderate",
        "Conservative", "Very Conservative"
    )
)

## pid3
ces$pid3 <- factor(ces$pid3,
    levels = 1:5,
    labels = c(
        "Democrat", "Republican", "Independent",
        "Other", "Not Sure"
    )
)

## pid7
ces$pid7 <- factor(ces$pid7,
    levels = 1:7,
    labels = c(
        "Strong Democrat", "Not very strong Democrat",
        "Lean Democract", "Independent", "Lean Republican",
        "Not very strong Republican", "Strong Republican"
    )
)

## medicare
ces$medicare <- factor(ces$medicare,
    levels = 1:2,
    labels = c("Support", "Oppose")
)

## ACA
ces$ACA <- factor(ces$ACA,
    levels = 1:2,
    labels = c("Support", "Oppose")
)

## abortion
ces$abortion <- factor(ces$abortion,
    levels = 1:2,
    labels = c("Support", "Oppose")
)

## EPA
ces$EPA <- factor(ces$EPA,
    levels = 1:2,
    labels = c("Support", "Oppose")
)

## Increase the number of police
ces$police_incr <- factor(ces$police_incr,
    levels = 1:2,
    labels = c("Support", "Oppose")
)

## Decrease the number of police
ces$police_decr <- factor(ces$police_decr,
    levels = 1:2,
    labels = c("Support", "Oppose")
)

## Trade Tariffs
ces$trade <- factor(ces$trade,
    levels = 1:2,
    labels = c("Support", "Oppose")
)

## Raise minimum wage
ces$minwage <- factor(ces$minwage,
    levels = 1:2,
    labels = c("Support", "Oppose")
)

## Work requirement for food stamps
ces$work_req <- factor(ces$work_req,
    levels = 1:2,
    labels = c("Support", "Oppose")
)






ces[1:2, ]
```



## Histogram
```{r}


ggplot(ces, aes(polengage)) +
    geom_histogram(color = "#a8a8a8", fill = "#000000", binwidth = 1) +
    labs(
        title = "Distribution of Political Engagment",
        ## caption = "Source: Gapminder dataset",
        x = "Political Activities",
        y = "Count"
    ) +
    theme_classic() +
    theme(
        plot.title = element_text(color = "#0099F8", size = 16, face = "bold"),
        plot.subtitle = element_text(size = 10, face = "bold"),
        plot.caption = element_text(face = "italic")
    )
```

```{r}
###Question 2 pt 1
## Education
Lm_educ <- lm(polengage ~ educ, data = ces)
Lm_educ

plot(polengage ~ educ, data = ces,
     xlab = "Education level", 
     ylab = "Number of Political Activities",
     main = "Impact of Education Level on Political Participation")


###Question 2 pt 2
ces$faminc <- (ifelse(ces$faminc == 2 | ces$faminc == 3, 2, ces$faminc))
ces$faminc <- (ifelse(ces$faminc == 4 | ces$faminc == 5, 3, ces$faminc))
ces$faminc <- (ifelse(ces$faminc == 6 | ces$faminc == 7, 4, ces$faminc))
ces$faminc <- (ifelse(ces$faminc == 8 | ces$faminc == 9, 5, ces$faminc))
ces$faminc <- (ifelse(ces$faminc == 10 | ces$faminc == 11, 6, ces$faminc))
ces$faminc <- (ifelse(ces$faminc == 12 | ces$faminc == 13, 7, ces$faminc))
ces$faminc <- (ifelse(ces$faminc == 14 | ces$faminc == 15, 8, ces$faminc))
ces$faminc <- (ifelse(ces$faminc == 16, 9, ces$faminc))

## Family income
ces$faminc <- factor(ces$faminc,
                     levels = 1:9,
                     labels = c(
                       "< 10,000", "10,000 - 29,999", "30,000 - 49,999", 
                       "50,000 - 69,999", "70,000 - 99,999", "100,000 - 149,999", "150,000 - 249,999", 
                       "250,000 - 499,999", "500,000 or more")
                     )


Lm_faminc <- lm(polengage ~ faminc, data = ces)
Lm_faminc

plot(polengage ~ faminc, data = ces,
     xlab = "Family Income", 
     ylab = "Number of Political Activities",
     main = "Impact of Family Income on Political Participation")   


### Question 2 pt 3 *It would be helpful here to turn the labels back to numerical###

plot(ces$faminc, ces$educ)
ces$educ <- as.numeric(ces$educ)
ces$faminc <- as.numeric(ces$faminc)
cor(ces$faminc, ces$educ, use = "complete.obs")


###Question 2 pt 4

##Trying to set party identification to part of a party or not##
ces$gender <- (ifelse(ces$gender == 1, 1, 0)) #1 if you are male, 0 if female
ces$race <- (ifelse(ces$race == 1, 1, 0)) #1 if you are white, 0 if non-white
ces$pid3 <- (ifelse(ces$pid3 == 1 | ces$pid3 == 2, 1, 0)) #1 if you identify as a member of a political party, 0 if not

Lm_our <- lm(polengage ~ gender + educ + race + faminc + pid3, data = ces)
summary(Lm_our)


###Question 3

Lm_abort <- lm(abortion ~ polengage, data = ces)
Lm_abort
summary (Lm_abort)

Lm_min <- lm(minwage ~ polengage, data=ces)
Lm_min
summary (Lm_min)
